# Re-process each dataset

# Load libraries

```{r, include=F}
library(Seurat)
library(tidyverse)
library(here)
library(reticulate)
library(MatrixExtra)

# test_scvi16_cuda113 is the author's conda env name, scrublet must be installed in your conda env: https://github.com/swolock/scrublet
use_condaenv('test_scvi16_cuda113', conda = "auto", required = TRUE)

```

# Set working directory

```{r,echo=F,eval=T,message=F,warning=F,error=F}

here::set_here()
print(paste("Current working directory:", here::here()))

```

# Load Integrated Data

```{r}
sample_name <- "Final_seurat_object.qs"
combined <- qs::qread(sample_name)
DefaultAssay(combined) <- "RNA"
combined

combine.list <- SplitObject(combined, split.by = "dataset")
```

```{r}
# Checks if the sparse matrix seems to be both normalized and log-transformed.
# Input: Seurat object.
# Output: TRUE if the matrix appears to be both normalized and log-transformed; otherwise FALSE.
is_normalized_and_log_transformed <- function(tmp_obj) {
  sparse_matrix <- GetAssayData(tmp_obj, slot = "data", assay = "RNA")
  
  # Randomly sample some values to check if they're not all integers
  sample_values <- sample(as.numeric(sparse_matrix), size = 1000, replace = TRUE)
  has_non_integers <- any(floor(sample_values) != sample_values)
  
  # Check if max value is less than a threshold, suggesting it's log-transformed
  max_val_below_threshold <- max(sparse_matrix) < 50
  
  # The matrix should not be all integers and max value should not exceed threshold
  # for it to be considered both normalized and log-transformed
  return(has_non_integers && max_val_below_threshold)
}

# Checks if the sparse matrix seems to be raw data.
# Input: Seurat object.
# Output: TRUE if the matrix seems to be raw data, otherwise FALSE.
is_raw_data <- function(tmp_obj) {
  sparse_matrix <- GetAssayData(tmp_obj, slot = "data", assay = "RNA")
  max_val <- max(sparse_matrix)
  sample_values <- sample(as.numeric(sparse_matrix), size = 1000, replace = TRUE)
  has_non_integers <- any(floor(sample_values) != sample_values)
  return(max_val > 50 && !has_non_integers)
}

# Runs Scrublet for doublet detection and adds results to Seurat object.
# Input: Seurat object with cells as rows and genes as columns.
# Output: Seurat object with a new metadata column named 'is_doublet_Scrublet'.
runScrublet <- function(tmp_obj) {
  sparse_matrix <- GetAssayData(tmp_obj, slot = "data", assay = "RNA")
  scr <- import("scrublet", convert = FALSE)
  scrub <- scr$Scrublet(as.matrix(t_shallow(sparse_matrix)), expected_doublet_rate = 0.05)
  predicted_doublets <- scrub$scrub_doublets()
  pred_score <- py_to_r(predicted_doublets[0])
  pred_doublet <- ifelse(pred_score > 0.3, T, F)
  tmp_obj <- AddMetaData(tmp_obj, pred_doublet, col.name = "is_doublet_Scrublet")
  return(tmp_obj)
}

# Re-runs preprocessing with quality filtering.
# Input: Original Seurat object with raw data.
# Output: New Seurat object, with quality filtering applied.
rerun_preprocessing <- function(tmp_obj) {
  original_n_cells <- ncol(tmp_obj)
  tmp_obj <- PercentageFeatureSet(tmp_obj, "^MT-", col.name = "percent.mito")
  if(ncol(tmp_obj) > 100) {
    tmp_obj <- suppressWarnings(runScrublet(tmp_obj))
  } else {
    tmp_obj <- AddMetaData(tmp_obj, F, col.name = "is_doublet_Scrublet")
  }
  n_doublets <- length(which(tmp_obj$is_doublet_Scrublet == T))
  tmp_obj <- AddMetaData(tmp_obj, n_doublets, col.name = "n_doublets")
  tmp_obj <- subset(tmp_obj, subset = percent.mito < 10 & nFeature_RNA > 200 & is_doublet_Scrublet == FALSE)
  new_n_cells <- ncol(tmp_obj)
  tmp_obj <- AddMetaData(tmp_obj, original_n_cells, col.name = "original_n_cells")
  tmp_obj <- AddMetaData(tmp_obj, new_n_cells, col.name = "new_n_cells")
  return(tmp_obj)
}

```


Processing dataset 1 (Adams_Kaminski_2020): Raw data = TRUE
Processing dataset 2 (Chen_Zhang_2020): Raw data = TRUE
Processing dataset 3 (Goveia_Carmeliet_2020): Raw data = TRUE
Processing dataset 4 (Guo_Zhang_2018): Raw data = TRUE
Processing dataset 5 (Habermann_Kropski_2020): Raw data = TRUE
Processing dataset 6 (He_Fan_2021): Raw data = TRUE
Processing dataset 7 (Lambrechts_Thienpont_2018_6149v1): Raw data = TRUE
Processing dataset 8 (Lambrechts_Thienpont_2018_6149v2): Raw data = TRUE
Processing dataset 9 (Lambrechts_Thienpont_2018_6653): Raw data = TRUE
Processing dataset 10 (Madissoon_Meyer_2020): Raw data = TRUE
Processing dataset 11 (Maier_Merad_2020): Raw data = TRUE
Processing dataset 12 (Reyfman_Misharin_2018): Raw data = TRUE
Processing dataset 13 (Travaglini_Krasnow_2020): Raw data = TRUE
Processing dataset 14 (UKIM-V): Raw data = TRUE
Processing dataset 15 (Vieira_Teichmann_2019): Raw data = TRUE
Processing dataset 16 (Wu_Zhou_2021): Raw data = TRUE
Processing dataset 17 (Zilionis_Klein_2019): Raw data = TRUE
Processing dataset 18 (Meyer_2021_5prime): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 19 (Schiller_2021): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 20 (Nawijn_2021): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 21 (Krasnow_2020): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 22 (Teichmann_Meyer_2019): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 23 (Meyer_2019): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 24 (Banovich_Kropski_2020): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 25 (Sun_2020_batch1): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 26 (Duong_lungMAP_unpubl): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 27 (Barbry_Leroy_2020): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 28 (Jain_Misharin_2021_10Xv2): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 29 (Lafyatis_Rojas_2019_10Xv2): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 30 (Tata_unpubl): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 31 (Sun_2020_batch3): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 32 (Misharin_2021): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 33 (MeyerNikolic_unpubl_UCL): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 34 (Meyer_2021_3prime): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 35 (Schultze_unpubl): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 36 (Misharin_Budinger_2018): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 37 (Schiller_2020): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 38 (Jain_Misharin_2021_10Xv1): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 39 (Sun_2020_batch2): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 40 (Lafyatis_Rojas_2019_10Xv1): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 41 (Sun_2020_batch4): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 42 (Eils_2020): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 43 (Shalek_2018): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 44 (Seibold_2020_10Xv3): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 45 (Seibold_2020_10Xv2): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 46 (Ahern_Zurke_2022): Raw data = TRUE
Processing dataset 47 (Data_Ji2020_Skin): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 48 (Data_Tirosh2016_Skin): Raw data = FALSE, Normalized and log-transformed = FALSE
Processing dataset 49 (Data_Chen2021_Prostate): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 50 (Data_Sun2021_Liver-Biliary): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 51 (Data_Bi2021_Kidney): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 52 (Data_Chen2020_Head and Neck): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 53 (Data_Lee2020_Colorectal): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 54 (Data_Li2017_Colorectal): Raw data = FALSE, Normalized and log-transformed = FALSE
Processing dataset 55 (Data_Gojo2020_Brain): Raw data = FALSE, Normalized and log-transformed = FALSE
Processing dataset 56 (Data_Neftel2019_Brain_10X): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 57 (Data_Neftel2019_Brain_SmartSeq2): Raw data = FALSE, Normalized and log-transformed = FALSE
Processing dataset 58 (Data_Venteicher2017_Brain): Raw data = FALSE, Normalized and log-transformed = FALSE
Processing dataset 59 (Data_Yuan2018_Brain): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 60 (Data_Jerby-Arnon2021_Sarcoma_SmartSeq2): Raw data = FALSE, Normalized and log-transformed = FALSE
Processing dataset 61 (Data_Qian2020_Breast): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 62 (Data_Galen2019_Hematologic): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 63 (Data_Wu2020_Hematologic_Group1): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 64 (Data_Wu2020_Hematologic_Group2): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 65 (Data_Wu2020_Hematologic_Group3): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 66 (Data_Wu2020_Hematologic_Group4): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 67 (Data_Qian2020_Lung): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 68 (Data_Xing2021_Lung): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 69 (Data_Dong2020_Neuroendocrine_Group1): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 70 (Data_Cui2021_Other-Models): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 71 (Data_Nath2021_Ovarian_10X): Raw data = FALSE, Normalized and log-transformed = TRUE
Processing dataset 72 (Data_Qian2020_Ovarian): Raw data = FALSE, Normalized and log-transformed = TRUE

# Process dataset

```{r}

new_list <- list()
# loop over each datasets
i=1
for (i in seq_along(combine.list)) {
  tmp_obj <- combine.list[[i]]
  
  check_raw_data <- suppressWarnings(is_raw_data(tmp_obj))
  check_normalized_and_log <-
    suppressWarnings(is_normalized_and_log_transformed(tmp_obj))
  
  # Print message
  status_message <- paste0("Processing dataset ",
                           i,
                           " (",
                           names(combine.list)[[i]],
                           "): Raw data = ",
                           check_raw_data)
  
  if (!check_raw_data) {
    status_message <- paste0(status_message,
                             ", Normalized and log-transformed = ",
                             check_normalized_and_log)
  }
  message(status_message)
  
  # Process labels
  if (check_raw_data) {
    tmp_obj <- rerun_preprocessing(tmp_obj)
    tmp_obj$data_label <- "raw_data"
    tmp_obj <- NormalizeData(tmp_obj)
  } else if (check_normalized_and_log) {
    tmp_obj$data_label <- "Normalized_Log"
  } else {
    tmp_obj$data_label <- "Normalized_nonLog"
    new_data_slot <- log1p(tmp_obj@assays$RNA@data)
    tmp_obj@assays$RNA@data <- new_data_slot
  }
  
  new_list[[i]] <- tmp_obj
  names(new_list)[i] <- names(combine.list[i])
}


qs::qsave(new_list, "new_seurat_obj_list.qsave")


```


# 2-Barplot: Number of normalized and unnormalized dataset

```{r}

dataset_labels <- sapply(new_list, function(x){
  return(as.character(x$data_label[1]))
})
df <- data.frame(labels = dataset_labels)
df_counts <- as.data.frame(table(dataset_labels))

p1 <- ggplot(df, aes(x = labels)) + 
  geom_bar(aes(fill = labels), show.legend = FALSE) +
  geom_text(data=df_counts, aes(x=dataset_labels, y=Freq, label=Freq), position=position_dodge(width=0.9), vjust=-0, size=6) + 
  xlab("Dataset Type") +
  ylab("Frequency") +
  ggtitle("Frequency of Dataset Types") +
  theme_classic()

print(p1)
```


# 3-Number and Barplot: How many cell are doublet; plot cell number ~ technologies

```{r}

j=2
doublet_tech <- data.frame()
for (j in seq_along(new_list)) {
  this_obj <- new_list[[j]]
  if ("is_doublet_Scrublet" %in% colnames(this_obj@meta.data)) {
    original_n_cells <- as.numeric(this_obj$original_n_cells[1])
    n_doublet <- as.numeric(this_obj$n_doublets[1])
    n_cells <- length(which(this_obj$is_doublet_Scrublet == FALSE))
    this_tech <- this_obj$technology[1]
    this_df <-
      data.frame(dataset = names(new_list[j]), n_doublet, n_cells, this_tech)
    doublet_tech <- rbind(doublet_tech, this_df)
  }
}

# Convert the matrix to data frame
doublet_tech_df <- as.data.frame(doublet_tech)
doublet_tech_df$n_doublet <- as.numeric(doublet_tech_df$n_doublet)
doublet_tech_df$n_cells <- as.numeric(doublet_tech_df$n_cells)
doublet_tech_df$this_tech <- as.character(doublet_tech_df$this_tech)

# Summarize data
summary_df <- doublet_tech_df %>%
  group_by(this_tech) %>%
  summarise(
    total_doublets = sum(n_doublet),
    total_cells = sum(n_cells)
  )

p2 <- ggplot(summary_df, aes(x = this_tech)) +
  geom_bar(aes(y = total_cells, fill = "Total Cells"), stat = "identity", position = "dodge") +
  geom_bar(aes(y = total_doublets, fill = "Doublets"), stat = "identity", position = "dodge") +
  geom_text(aes(y = total_cells, label = total_cells), position=position_dodge(width=0.9), vjust=0, size=6) +
  geom_text(aes(y = total_doublets, label = total_doublets), position=position_dodge(width=0.9), vjust=0.5, size=6) +
  ylab("Count") +
  labs(fill = "Type") +
  theme_classic()

print(p2)
```

# 4-Number: how many cell are removed in quality control step.

```{r}

j=1
total_removed_cells <- data.frame()
for (j in seq_along(new_list)) {
  this_obj <- new_list[[j]]
  if("original_n_cells" %in% colnames(this_obj@meta.data)) {
    original_n_cells <- as.numeric(this_obj$original_n_cells[1])
  new_n_cells <- as.numeric(this_obj$new_n_cells[1])
  removed_cells <- original_n_cells - new_n_cells
  this_df <- data.frame(dataset = names(new_list[j]), original_n_cells, new_n_cells, removed_cells)
  total_removed_cells <- rbind(total_removed_cells, this_df)
  }
}

total_removed_cells
write.csv(total_removed_cells, "total_removed_cells.csv")
```



```{r}


```

```{r}


```

```{r}


```

```{r}


```

```{r}


```

```{r}


```

```{r}


```

```{r}


```

```{r}


```

```{r}


```

```{r}
sessionInfo()

```

